A Flexible Space-time Tradeoff on Hybrid Index with Bicriteria Optimization

### 2016-08-07 manuscript.tex
alternate some figures
more fixups.

\section{Introduction}\label{sec:intro}
Indexing is essential to efficiently manage growing data and perform quick queries in many information retrieval systems.
The use of inverted Indexes in web search engine is probably one of the most successful examples in IR.
Despite of the fact that it is a relatively old yet simple data structure, inverted index is extensible with supplementary information to answer different queries within much less space occupied.
Compression has been one of the most active field of inverted index research in the last decades, for it not only reduces the space occupancy, but also accelerates data transfer in I/O.
See \cite{catena2014inverted,trotman2014compression,zobel2006inverted} for a detailed survey.

Typical inverted index is composed of two parts: \textit{lexicon} and \textit{posting lists}.
The inverted index compression we are talking about is usually concentrated in the compression of posting lists.
However, there exists a divergence between compressed size and decompression time,
since pursuing the minimization of size will undoubtedly impair decoding.
Encoders with different space-time treadoffs keep coming out to settle for requirements of various user scenarios \cite{lemire2015decoding,ottaviano2015optimal,ottaviano2014partitioned,petri2014score,silvestri2010vsencoding}.
Among these techniques, splitting posting list into \textit{blocks} (or \textit{chunks}) has been demonstrate to be an efficient way and widely adopted by many encoders.
For the reason that by keeping blocks separately accessible we can expect faster decoding speed and hierarchical structure.
In addition, variable-sized blocks can utilize the clustering property of list to beat the entropy \cite{moffat2000binary,silvestri2010vsencoding}.
Nonetheless, designers of information systems have to carefully choose one of these encoders to satisfy different IR tasks: whether to achieve optimal compression ratio/decoding speed, sacrificing the other one; or to balance them by adopting one proper encoder with finely tuned space-time tradeoff.

@@

This idea is first adopted in \cite{yan2009inverted}.
However, the procedure of choosing encoders works at a posting list granularity, which is too coarse to obtain a flexible space-time tradeoff.
They also fail to explore complexity of the algorithm as it works on a empirical foundation.
Giuseppe et al. \cite{ottaviano2015optimal} explicitly define the Space-Constrained Expected Time Minimization problem: given a dataset $ D $ and an upper bound $ \cal S $ on its compressed size, the goal is to choose one proper encoder from an encoder set $ \cal E $ for each block of the index.
Thus minimizing the expected query processing time, provided that it can be compressed within $ \cal S $ space.
They then propose a linear time solution for this problem, which is quite straight forward.
Under two predefined dominance criteria, their algorithm works by enumerating all the advisable compression methods, then sorting them in an non increasing order by the time/space gain ratio of each advisable encoder on each block.
After this, a sequential scan and replace is able to find all the optional solutions for any possible budget constraint $ \cal S $.
Given the fact that each posting list is divided into fixed-sized blocks, whose total number is $ n $, and a set of $ k $ encoders, the algorithm finishes in $ \Theta(nk\log(nk)) $ time and $ O(nk) $ space.

@@

We argue that global ordering of the compression options for all the blocks is not a good idea, since it will lead to the above three shortcomings.
In this paper, we try to solve this problem from another perspective.
The main contributions are listed below:
\begin{enumerate}
	\item We introduce the conception of \textit{Bicriteria Compression} \cite{farruggia2014bicriteria} to extend the Space-Constrained Expected Time Minimization problem.
	Except the original one, it models another dual problem: we could exchange the role of space and time by asking for the minimal compressed size, provided an upper bound $ T $ in its average query processing time or decompression time.
	By solving the \textit{Bicriteria Compression} problem we can expect the compression method to be more universal for different scenarios.
	\item We reduce the problem into a Resource Constrained Shortest Path problem \cite{mehlhorn2000resource} over a weighted DAG $ \cal G $, whose edges contain two kinds of penalties: space cost and time weight.
	Thus, the problem can be formalized using Lagrangian relaxation, we use the algorithm introduced in \cite{farruggia2014bicriteria} and \cite{handler1980dual} to settle in $ O(n\log n) $ time and $ O(n) $ space, while keeping the same additive approximation guarantee.
	\item Different from \cite{ottaviano2015optimal} which is a greedy algorithm, our method amortizes the expected budget over each posting list in a more balanced way.
	Thus we could expect for a better average performance.
	While keeping the same space and time complexity, our method is more efficient and needs less runtime memory in practice without global sorting.
	\item We execute a preliminary set of experiments on two realistic dataset, GOV2 and Common Crawl, with AOL query log.
	Analysis shows that our method is able to obtain flexible tradeoffs under bicriteria compression.
	Its compressed size and query processing time is very competitive to the hybrid index in \cite{ottaviano2015optimal}, however, with more detailed description and more strict theoretic foundation.
\end{enumerate}

@@

\subsection{Closing the Gap}
Upon termination of the iteration we have two paths $ (\pi_l,$ $\pi_r) $ which specify a lower bound and an upper bound of the minimal value $ f(\pi^\star) $.
From Equation~\eqref{equ: maximum}, we know the lower bound is the intersection point, namely $ LB = L(\mu^\star) $; also, any path $ \pi $ with $ g(\pi) \leqslant 0 $ generated during the iteration is a feasible solution for RCSP, and the best value for upper bound is $ f(\pi_r) $ from the last path $ \pi_r $.
Thus we have the bound $ L(\mu^\star) \leqslant f(\pi^\star) \leqslant f(\pi_r) $, which is also shown in Figure~\ref{fig: intersection}.

@@

\subsection{Indexing Performance}
\paragraph*{Construction time}
We have analyzed our algorithm in detail in the previous section.
Its complexity is $ O(\tilde{n} $$ \log(n $ $ s_{costs} $$ t_{costs})) $ in time and $ O(\tilde{n}) $ in space, almost equaled by \textit{hybrid} which is $ O(\tilde{n} $$ \log(\tilde{n})) $ in time and $ O(\tilde{n}) $ in space.
Both \textit{bicriteria} and \textit{hybrid} concatenate all the posting lists into one to arrange encodings of each block, and are delayed by one time-consuming phase that cannot be paralleled, for \textit{bicriteria} is the iteration when solving the dual problem, for \textit{hybrid} is the sorting of $ \lambda_{i,j} $.

@@

As shown in Figure~\ref{fig: cc_S}, by linearly relaxing the budget in space, \textit{bicriteria} trends to choose more fast encoders.
At first more than 90\% blocks are compressed using Interpolative, then more PFD-compressed blocks are involved and last nearly all of them are compressed using Varint-G8IU.
The result agrees with that of \cite{ottaviano2015optimal}, however, their space-time curve is more smoothing because when the space budget changes, \textit{hybrid} executes a sequential scan to replace encoders right starting from the last block it stopped.
\textit{Bicriteria} works in a more \textbf{balanced} way, the budget is amortized to all the blocks since it relies on path finding rather than sorting, while \textit{hybrid} is a \textbf{greedy} algorithm, it always prefer blocks which are more frequently accessed.
\begin{figure}
	\centering
	\subfloat[linearly varying space budget $ S $ from 0 to 1 by 0.1]{
		\label{fig: cc_S}
		\includegraphics[width=1.0\linewidth]{cc_S}
	}\\
	\subfloat[linearly varying time budget $ T $ from 1 to 0 by 0.1]{
		\label{fig: cc_T}
		\includegraphics[width=1.0\linewidth]{cc_T}
	}
	\caption{Percentage of encoders used when compressing blocks under different space/time budgets for Common Crawl, the abscissa represents compressed size in GB.}
	\label{fig: percentage}
\end{figure}

>>>>

\section{Introduction}\label{sec:intro}
Indexing is essential to efficiently managing growing data and performing quick queries in many information retrieval systems.
The use of inverted Indexes in web search engine is probably one of the most successful examples in IR.
Despite of the fact that it is a relatively old yet simple data structure, inverted index is extensible with supplementary information to answer different queries within much less space occupied.
Compression has been one of the most active field of inverted index research in the last decades, for it not only reduces the space occupancy, but also accelerates data transfer in I/O.
See \cite{catena2014inverted,trotman2014compression,zobel2006inverted} for a detailed survey.

Typical inverted index is composed of two parts: \textit{lexicon} and \textit{posting lists}.
The inverted index compression we are talking about is usually concentrated in the compression of posting lists.
However, there exists a divergence between compressed size and decompression time,
since pursuing the minimization of size will undoubtedly impair decoding.
Encoders with different space-time treadoffs keep coming out to settle for requirements of various user scenarios \cite{lemire2015decoding,ottaviano2015optimal,ottaviano2014partitioned,petri2014score,silvestri2010vsencoding}.
Among these techniques, splitting posting list into \textit{blocks} (or \textit{chunks}) has been demonstrated to be an efficient way and is widely adopted by many encoders.
For the reason that by keeping blocks separately accessible we can expect faster decoding speed and hierarchical structure.
In addition, variable-sized blocks can utilize the clustering property of list to beat the entropy \cite{moffat2000binary,silvestri2010vsencoding}.
Nonetheless, designers of information systems have to carefully choose one of these encoders to satisfy different IR tasks: whether to achieve optimal compression ratio/decoding speed, sacrificing the other one; or to balance them by adopting one proper encoder with finely tuned space-time tradeoff.

@@

This idea is first adopted in \cite{yan2009inverted}.
However, the procedure of choosing encoders works at a posting list granularity, which is too coarse to obtain a flexible space-time tradeoff.
They also fail to explore complexity of the algorithm as it works on a empirical foundation.
Giuseppe et al. \cite{ottaviano2015optimal} explicitly define the Space-Constrained Expected Time Minimization problem: given a dataset $ D $ and an upper bound $ \cal S $ on its compressed size, the goal is to choose one proper encoder from an encoder set $ \cal E $ for each block of the index.
Thus minimizing the expected query processing time, provided that it can be compressed within $ \cal S $ space.
They then propose a linear time solution for this problem, which is quite straight forward.
Under two predefined dominance criteria, their algorithm works by enumerating all the advisable compression methods, then sorting them in an non-increasing order by the time/space gain ratio of each advisable encoder on each block.
After this, a sequential scan and replace is able to find all the optional solutions for any possible budget constraint $ \cal S $.
Given the fact that each posting list is divided into fixed-sized blocks, whose total number is $ n $, and a set of $ k $ encoders, the algorithm finishes in $ \Theta(nk\log(nk)) $ time and $ O(nk) $ space.

@@

We argue that global ordering of the compression options for all the blocks is not a good idea, since it will lead to the above three shortcomings.
In this paper, we try to solve this problem from another perspective.
The main contributions are listed below:
\begin{enumerate}
	\item We introduce the conception of \textit{Bicriteria Compression} \cite{farruggia2014bicriteria} to extend the Space-Constrained Expected Time Minimization problem.
	Except the original one, it models another dual problem: we could exchange the role of space and time by asking for the minimal compressed size, provided an upper bound $ T $ in its average query processing time or decompression time.
	By solving the \textit{Bicriteria Compression} problem we can expect the compression method to be more universal for different scenarios.
	\item We reduce the problem into a Resource Constrained Shortest Path problem \cite{mehlhorn2000resource} over a weighted DAG $ \cal G $, whose edges contain two kinds of penalties: space cost and time weight.
	Thus, the problem can be formalized using Lagrangian relaxation, we use the algorithm introduced in \cite{farruggia2014bicriteria} and \cite{handler1980dual} to settle in $ O(n\log n) $ time and $ O(n) $ space, while keeping the same additive approximation guarantee.
	\item Different from \cite{ottaviano2015optimal} which is a greedy algorithm, our method amortizes the expected budget over each posting list in a more balanced way.
	Thus we could expect for a better average performance.
	While keeping the same space and time complexity, our method is more efficient and needs less runtime memory in practice without global sorting.
	\item We execute a detailed experimental evaluation on two realistic dataset, GOV2 and Common Crawl, with AOL query log.
	Analysis shows that our method is able to obtain flexible tradeoffs under bicriteria compression.
	Its compressed size and query processing time are very competitive to the hybrid index in \cite{ottaviano2015optimal}, however, with more detailed description and more strict theoretic foundation.
\end{enumerate}

@@

\subsection{Closing the Gap}
Upon termination of the iteration we have two paths $ (\pi_l, $ $ \pi_r) $ which specify a lower bound and an upper bound of the minimal value $ f(\pi^\star) $.
From Equation~\eqref{equ: maximum}, we know the lower bound is the intersection point, namely $ LB = L(\mu^\star) $; also, any path $ \pi $ with $ g(\pi) \leqslant 0 $ generated during the iteration is a feasible solution for RCSP, and the best value for upper bound is $ f(\pi_r) $ from the last path $ \pi_r $.
Thus we have the bound $ L(\mu^\star) \leqslant f(\pi^\star) \leqslant f(\pi_r) $, which is also shown in Figure~\ref{fig: intersection}.

@@

\subsection{Indexing Performance}
\paragraph*{Construction time}
We have analyzed our algorithm in detail in the previous section.
Its complexity is $ O(\tilde{n} \log(n $ $ s_{costs} t_{costs})) $ in time and $ O(\tilde{n}) $ in space, almost equaled by \textit{hybrid} which is $ O(\tilde{n}\log(\tilde{n})) $ in time and $ O(\tilde{n}) $ in space.
Both \textit{bicriteria} and \textit{hybrid} concatenate all the posting lists into one to arrange encodings of each block, and are delayed by one time-consuming phase that cannot be paralleled, for \textit{bicriteria} is the iteration when solving the dual problem, for \textit{hybrid} is the sorting of $ \lambda_{i,j} $.

@@

As shown in Figure~\ref{fig: cc_S}, by linearly relaxing the budget in space, \textit{bicriteria} trends to choose more fast encoders.
At first more than 90\% blocks are compressed using Interpolative, then more PFD-compressed blocks are involved and last nearly all of them are compressed using Varint-G8IU.
The result agrees with that of \cite{ottaviano2015optimal}, however, their space-time curve is more smoothing because when the space budget changes, \textit{hybrid} executes a sequential scan to replace encoders right starting from the last block it stopped.
\textit{Bicriteria} works in a more \textbf{balanced} way, the budget is amortized to all the blocks since it relies on path finding rather than sorting, while \textit{hybrid} is a \textbf{greedy} algorithm, it always prefer blocks which are more frequently accessed.
\begin{figure}
	\centering
	\subfloat[linearly varying space budget $ S $ from 0 to 1 by 0.1]{
		\label{fig: cc_S}
		\includegraphics[width=0.95\linewidth]{S}
	}\\
	\subfloat[linearly varying time budget $ T $ from 1 to 0 by 0.1]{
		\label{fig: cc_T}
		\includegraphics[width=0.95\linewidth]{T}
	}
	\caption{Percentage of encoders used when compressing blocks under different space/time budgets for Common Crawl, the abscissa represents compressed size in GB.}
	\label{fig: percentage}
\end{figure}
